*Prochlorococcus marinus* responses to light and oxygen

### 2024-03-03 Code/Manuscript_O2.Rmd
Fixed reference to FtsH 2 and 3 to FtsH 1 and 2

Figure \@ref(fig:ProchlorococcusEnzymeKms) shows the measured or inferred K~M~ for [O~2~] for genes encoding enzymes [@omarAnnotationGenesEncoding2023] for  *P. marinus* strains from clades HLI, LLI, LLIII and LLIV. XXXX CITATION  to Table once added XXXX. MED4 increases expression of alternative oxidase ('ubiquinol oxidase (non electrogenic)')  to cope with changes in light [@bergResponsesPsbAHli2011], by dissipating electrons from the inter-system transport chain. The approximate K~M~ for [O~2~] of ~ 25 µM for ubiquinol oxidase (non electrogenic) is comparable to the lower limit for growth of MED4 in our experiments.  We suggest that dependence upon this enzyme excludes MED4 from low oxygen zones.  The genome scan shows SS120 and MIT9313 lack this gene, and therefore, lack this oxygen-dependent path to cope with changing excitation.  Conversely, a gene encoding (S)-2-hydroxy-acid oxidase is encoded in the MIT9313 genome.  (S)-2-hydroxy-acid oxidase catalyzes the reaction of 2-hydroxy acid with O~2~ to produce toxic H~2~O~2~ [@cunaneCrystalStructureAnalysis2005]. (S)-2-hydroxy-acid oxidase has an approximate K~M~ for [O~2~] of ~ 250 µM, and produces H~2~O~2~, so growth at lower [O~2~] may protect MIT9313 from auto-intoxication from production of H~2~O~2~. We hypothesize that under 250 µM O~2~ and higher blue light, *P. marinus* MIT9313 suffered photoinhibition, resulting from the inactivation of PSII caused by the production of the reactive oxygen species, hydrogen peroxide. This photoinhibition is compounded by the limited inducible repair mechanism for PSII, due to the absence of FtsH 2 and 3 XXXXCHECK TERMINOLOGYXXXX expression in *P. marinus* MIT9313 [@bonisteelStrainSpecificDifferences2018].  We hypothesize that under the conditions of our high light and 2.5 µM or 25 µM O~2~ experiments, the activity of the (S)-2-hydroxy-acid oxidase enzyme is suppressed. As a result, the catalyzed production of hydrogen peroxide is inhibited, leading to less PSII damage, allowing MIT9313 to avoid photoinhibition and circumvent its limitations on PSII repair to exploit higher light.  Figure \@ref(fig:ProchlorococcusEnzymeKms) also shows that *P. marinus* SS120 is the only tested ecotype to lack the pyridoxal 5'-phosphate synthase enzyme. The pyridoxal 5'-phosphate synthase enzyme is an important cofactor in the biosynthesis of vitamin B~6~ [@sangIdentificationPyridoxinePyridoxamine2007]. Vitamin B~6~ is a potential antioxidant and can effectively quench singlet oxygen [@bilskiVitaminB6Pyridoxine2000]. The absence of the pyridoxal 5'-phosphate synthase enzyme may explain why *P. marinus* SS120 does not grow as well as *P. marinus* MIT9313, when exposed to high light stress under 25 µM O~2~ and not at all under 2.5 µM O~2~.

@@

*Prochlorococcus marinus* MIT9313, a LLIV clade representative, shows the potential to inhabit future warmer temperate zones year-round, as it grows under a 4 h photoperiod, expected in winter or at light-attenuated depths. MIT9313 demonstrates an unexpected tolerance to higher light levels and cumulative diel PUR,  but only under low oxygen conditions of 25 µM and 2.5 µM (Figure \@ref(fig:MIT9313PAR)). The ability to utilize higher PAR and higher cumulative diel PUR under 25 µM or 2.5 µM O~2~ enables MIT9313 to grow in OMZ, even at depths closer to the surface. MIT9313 carries a gene encoding (S)-2-hydroxy-acid oxidase [@cunaneCrystalStructureAnalysis2005],  with a K~M~ for [O~2~] of ~ 250 µM (Figure \@ref(fig:ProchlorococcusEnzymeKms)), which produces H~2~O~2~.  Growth at lower [O~2~] may protect MIT9313 from auto-intoxication from production of H~2~O~2~. We hypothesize that under 250 µM O~2~ and higher blue light, *P. marinus* MIT9313 suffers photoinhibition, resulting from the inactivation of PSII caused by the production of H~2~O~2~. This photoinhibition is compounded by the limited inducible repair for PSII, due to the absence of FtsH 2 and 3  expression in *P. marinus* MIT9313 [@bonisteelStrainSpecificDifferences2018].  MIT9313 shows remarkable ability to thrive under very low [O~2~], potentially allowing it to expand into broader ecological niches. These results are supported by Figure \@ref(fig:CladeProchloroPhotosynthDepthO2) showing PSI protein subunits annotated as derived from MIT9313 detected at depths \> 120 m, along with PSII subunits at depths from 50 m to 200 m in regions where O~2~ was 15 µM.

>>>>

Figure \@ref(fig:ProchlorococcusEnzymeKms) shows the measured or inferred K~M~ for [O~2~] for genes encoding enzymes [@omarAnnotationGenesEncoding2023] for  *P. marinus* strains from clades HLI, LLI, LLIII and LLIV. XXXX CITATION  to Table once added XXXX. MED4 increases expression of alternative oxidase ('ubiquinol oxidase (non electrogenic)')  to cope with changes in light [@bergResponsesPsbAHli2011], by dissipating electrons from the inter-system transport chain. The approximate K~M~ for [O~2~] of ~ 25 µM for ubiquinol oxidase (non electrogenic) (Figure \@ref(fig:ProchlorococcusEnzymeKms)) is comparable to the lower limit for growth of MED4 in our experiments (Figure \@ref(fig:MED4GAM)).  We suggest that dependence upon this enzyme excludes MED4 from low oxygen zones.  The genome scan shows SS120 and MIT9313 lack this gene (Figure \@ref(fig:ProchlorococcusEnzymeKms)), and therefore, lack this oxygen-dependent path to cope with changing excitation.  Conversely, a gene encoding (S)-2-hydroxy-acid oxidase is encoded in the MIT9313 genome (Figure \@ref(fig:ProchlorococcusEnzymeKms)).  (S)-2-hydroxy-acid oxidase catalyzes the reaction of 2-hydroxy acid with O~2~ to produce toxic H~2~O~2~ [@cunaneCrystalStructureAnalysis2005]. (S)-2-hydroxy-acid oxidase has an approximate K~M~ for [O~2~] of ~ 250 µM, and produces H~2~O~2~, so growth at lower [O~2~] may protect MIT9313 from auto-intoxication from production of H~2~O~2~. We hypothesize that under 250 µM O~2~ and higher blue light, *P. marinus* MIT9313 suffered photoinhibition, resulting from the inactivation of PSII caused by the production of the reactive oxygen species, hydrogen peroxide. This photoinhibition is compounded by the limited inducible repair mechanism for PSII, due to the absence of FtsH 1 and 2 XXXXCHECK TERMINOLOGYXXXX expression in *P. marinus* MIT9313 [@bonisteelStrainSpecificDifferences2018].  We hypothesize that under the conditions of our high light and 2.5 µM or 25 µM O~2~ experiments, the activity of the (S)-2-hydroxy-acid oxidase enzyme is suppressed. As a result, the catalyzed production of hydrogen peroxide is inhibited, leading to less PSII damage, allowing MIT9313 to avoid photoinhibition and circumvent its limitations on PSII repair to exploit higher light.  Figure \@ref(fig:ProchlorococcusEnzymeKms) also shows that *P. marinus* SS120 is the only tested ecotype to lack the pyridoxal 5'-phosphate synthase enzyme. The pyridoxal 5'-phosphate synthase enzyme is an important cofactor in the biosynthesis of vitamin B~6~ [@sangIdentificationPyridoxinePyridoxamine2007]. Vitamin B~6~ is a potential antioxidant and can effectively quench singlet oxygen [@bilskiVitaminB6Pyridoxine2000]. The absence of the pyridoxal 5'-phosphate synthase enzyme may explain why *P. marinus* SS120 does not grow as well as *P. marinus* MIT9313, when exposed to high light stress under 25 µM O~2~ and not at all under 2.5 µM O~2~.

@@

*Prochlorococcus marinus* MIT9313, a LLIV clade representative, shows the potential to inhabit future warmer temperate zones year-round, as it grows under a 4 h photoperiod, expected in winter or at light-attenuated depths. MIT9313 demonstrates an unexpected tolerance to higher light levels and cumulative diel PUR,  but only under low oxygen conditions of 25 µM and 2.5 µM (Figure \@ref(fig:MIT9313PAR)). The ability to utilize higher PAR and higher cumulative diel PUR under 25 µM or 2.5 µM O~2~ enables MIT9313 to grow in OMZ, even at depths closer to the surface. MIT9313 carries a gene encoding (S)-2-hydroxy-acid oxidase [@cunaneCrystalStructureAnalysis2005],  with a K~M~ for [O~2~] of ~ 250 µM (Figure \@ref(fig:ProchlorococcusEnzymeKms)), which produces H~2~O~2~.  Growth at lower [O~2~] may protect MIT9313 from auto-intoxication from production of H~2~O~2~. We hypothesize that under 250 µM O~2~ and higher blue light, *P. marinus* MIT9313 suffers photoinhibition, resulting from the inactivation of PSII caused by the production of H~2~O~2~. This photoinhibition is compounded by the limited inducible repair for PSII, due to the absence of FtsH 1 and 2  expression in *P. marinus* MIT9313 [@bonisteelStrainSpecificDifferences2018].  MIT9313 shows remarkable ability to thrive under very low [O~2~], potentially allowing it to expand into broader ecological niches. These results are supported by Figure \@ref(fig:CladeProchloroPhotosynthDepthO2) showing PSI protein subunits annotated as derived from MIT9313 detected at depths \> 120 m, along with PSII subunits at depths from 50 m to 200 m in regions where O~2~ was 15 µM.
